---
title: "Believers vs NonBelievers"
author: "Elizabeth Hiroyasu"
date: "October 14, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
library(sandwich)
library(car)
library(psych)

grizz <- read.csv("GRIZZ_WT_VALUES.csv", header = T)
grizz<-grizz[,-1:-3]
```


##Believers vs NonBelievers

```{r, echo=FALSE, warning=FALSE}
ggplot(grizz[!is.na(grizz$species_griz),], aes(x=factor(species_griz))) + 
  geom_bar(aes(y=(..count..)/sum(..count..)), fill = c("darkred", "darkgreen", "darkred")) +
  labs(x = "Response", y="Percent", title = "Do Grizzly Bears Exist in California") + 
  scale_y_continuous(labels=percent) +
  scale_x_discrete(breaks = 1:3, labels = c("Yes", "No", "Don't Know")) + 
  theme(panel.border = element_rect(linetype = "solid", colour = "black", fill = NA))


names(grizz)[46:60]<-c("survival", "forests", "SafetyThreat", "LivelihoodThreat", "Tourism",
                       "Feds", "ThreatPP", "OtherSpecies", "Extinction", "LocalControl", 
                       "Ranchers", "OutdoorRec", "Ag", "Urban", "Rural")

believers<-(grizz[which(grizz$species_griz==1),])
highknowledge<-(grizz[which(grizz$species_griz==2),])
dontknow<-(grizz[which(grizz$species_griz==3),])
lowknowledge<-rbind(believers, dontknow)
```

##Different species
```{r}
species_df<-subset(grizz, select=species_griz:species_bbear)
names(species_df)<-c("Grizzly Bear", "Bald Eagle", "Bison", "Wolf", "Black Bear")

library(reshape2) # for melt()
species_df <- melt(species_df)

ggplot(species_df)  + 
 geom_bar(aes(as.factor(value), fill=as.factor(value))) + 
  labs(x="Response", y="Frequency", title="Do the following species exist in California?")+
  scale_x_discrete(breaks=1:4, labels=c("Yes", "No", "Don't Know", "NA"))+
  guides(fill=FALSE)+
  facet_wrap(~ variable)
```


##Factor analysis
#Believers
Incorrectly identified grizzly bears as currently existing in CA.
```{r, echo=FALSE}
believe_comp<-factanal(believers[,c(46:47, 50:60)], 4, scores="Bartlett")
believe_comp

#Plotting the relationship of the factors
fa.diagram(believe_comp$loadings, main="Believers Factor Analysis")

believers<-cbind(believers, believe_comp$scores)
names(believers)[names(believers)=="Factor2"]<-"EcologicalImpacts"
names(believers)[names(believers)=="Factor3"]<-"Benefits"
write.csv(believers,"believers_factors.csv")
```

#High Knowledge
Correctly believed that grizzly bears do not currently exist in CA.
```{r, echo=FALSE}
highknowledge_comp<-factanal(highknowledge[,c(46:47, 50:60)],4, scores="Bartlett")
#highknowledge_comp

#Plotting the relationship of the factors
fa.diagram(highknowledge_comp$loadings, main="High Knowledge Factor Analysis")

highknowledge<-cbind(highknowledge, highknowledge_comp$scores)
names(highknowledge)[names(highknowledge)=="Factor1"]<-"Benefits"
names(highknowledge)[names(highknowledge)=="Factor2"]<-"EcologicalImpacts"

write.csv(highknowledge, "nonbelievers_factors.csv")
```

#Don't Know
Did not know if grizzly bears existed in CA
```{r, echo=FALSE}
dontknow_comp<-factanal(dontknow[,c(46:47, 50:60)], 4, scores="Bartlett")
#dontknow_comp

#Plotting the relationship of the factors
fa.diagram(dontknow_comp$loadings, main="Don't Know Factor Analysis")

dontknow<-cbind(dontknow, dontknow_comp$scores)
names(dontknow)[names(dontknow)=="Factor1"]<-"Benefits"
names(dontknow)[names(dontknow)=="Factor2"]<-"EcologicalImpacts"
write.csv(dontknow, "dontknow_factors.csv")
```

#Low Knowledge
Pooled Don't know and respondents who incorrectly answered the question about whether grizzly bears currently exist in CA.
```{r, echo=FALSE}
lowknowledge_comp<-factanal(lowknowledge[,c(46:47, 50:60)], 4, scores="Bartlett")
#lowknowledge_comp

#Plotting the relationship of the factors
fa.diagram(lowknowledge_comp$loadings, main="Low Knowledge Factor Analysis")

lowknowledge<-cbind(lowknowledge, lowknowledge_comp$scores)
names(lowknowledge)[names(lowknowledge)=="Factor1"]<-"EcologicalImpacts"
names(lowknowledge)[names(lowknowledge)=="Factor2"]<-"Benefits"
write.csv(lowknowledge, "lowknowledge_factors.csv")
```

##Testing differences between High knowledge and Low Knowledge Respondents
```{r, echo=FALSE}

highknowledge$altruism_weights<-highknowledge$altruism*highknowledge$weights
highknowledge$egoism_weights<-highknowledge$egoism*highknowledge$weights
highknowledge$biospherism_weights<-highknowledge$biospherism*highknowledge$weights

lowknowledge$altruism_weights<-lowknowledge$altruism*lowknowledge$weights
lowknowledge$egoism_weights<-lowknowledge$egoism*lowknowledge$weights
lowknowledge$biospherism_weights<-lowknowledge$biospherism*lowknowledge$weights

t.test(highknowledge$altruism_weights, lowknowledge$altruism_weights)
t.test(highknowledge$egoism_weights, lowknowledge$egoism_weights)
t.test(highknowledge$biospherism_weights, lowknowledge$biospherism_weights)

boxplot(highknowledge$altruism_weights, lowknowledge$altruism_weights,
        highknowledge$egoism_weights, lowknowledge$egoism_weights,
        highknowledge$biospherism_weights, lowknowledge$biospherism_weights,
        main="Difference of means of values for high and low knowledge respondents",
        at=c(1,2, 4,5, 7,8), names=c("high\n alt", "low\n alt", "high\n ego", 
                                     "low\n ego", "high\n bio", "low\n 
                                     bio"),
        col=c("firebrick4", "darkgreen"))

```