---
title: "Grizzly Attitudes"
author: "Elizabeth Hiroyasu"
date: "September 12, 2017"
output: html_document
---

Factor analysis conducted to assess how people think about grizzly bear reintroduction and what types of variables those attitudes map onto.
```{r setup, include=FALSE}
library(survey)
library(weights)
library(anesrake)
library(lmtest)
library(sandwich)
library(car)
library(psych)

grizz <- read.csv("GRIZZLY_DATA_USE.csv", header = T)
```

##Knowledge
```{r, echo=FALSE}
ggplot(grizz) + 
  geom_bar(aes(species_griz),bins = 3, width=1,  fill = c("darkgreen", "firebrick4", "sienna4"), colour = "black") +
  labs(x = "Response", y = "Frequency",title = "Do Grizzly Bears Exist in California") +
  scale_x_continuous(breaks = 1:3, labels = c("Yes", "No", "Don't Know")) + 
  theme(text = element_text(family = "Trebuchet MS"),
        plot.title = element_text(size = 14, face = "bold"), 
        axis.title.x = element_text(size = 10, face = "bold"),
        axis.title.y = element_text(size = 10, face = "bold"),
        panel.border = element_rect(linetype = "solid", colour = "black", fill = NA))

```

##Weighting
For this survey we oversampled rural residents, so we have to weight the data to account for this:
```{r, echo=FALSE}
# Create dummy for highet_degree and urban_rural so I can weight according to whether 
# someone is (not) rural and whether they (don't) have college degree. This is jsut a 
# formality type thing to make the weighting process go more smoothly. 

grizz$college <- grizz$highest_degree>=5 
grizz$rural <- grizz$urban_rural==3

# Here I isolate percent of respondents falling into each category. This is so I have 
# a rough estimate of distribution of respondents. 

wpct(grizz$gender)
wpct(grizz$urban_rural)
wpct(grizz$latino)
wpct(grizz$highest_degree)
wpct(grizz$college)
wpct(grizz$rural)

# wpct produces a table of the proportion of data in each category for any variable.

# Here I use US Census data to make variables with distribution of each data for all of 
# California. 

gender <- c(.50, .50)
rural <- c(.13, .87) # Rural; non-rural.
latino <- c(.39, .61) # Latino; not-latino.
college <- c(.32, .68) # BA/higher; no college degree.  

# Here I put the census variables into a list and name them. This is so they can be called 
# upon later and used for weighting. Essentially, they are population-level estiamtes for 
# all of California and are what we weight to. 

targets <- list(gender, rural, latino, college)
names(targets) <- c("gender", "rural", "latino", "college")

# Here I create unique ids for all complete observations in the dataframe, and then determine
# which variables should be used to weight based on the ones I provide in "target" list. 

grizz$caseid <- 1:length(grizz$good_complete) # Create unique id for each complete obs. 
anesrakefinder(targets, grizz, choosemethod="total") 

# Here I rake weights data. This process determines how variables should be weighted and then rakes the
# data to develop weights for the variables so that they match the target values I provided. 

newgrizz <- anesrake(targets, grizz, caseid = grizz$caseid,
                     verbose=F, cap=5, choosemethod = "total",
                     type="pctlim", pctlim=.05, nlim=5, iterate=T, force1=T)

# Here I create a new dataframe with weighted data #

wt_grizz_df <- svydesign(ids = ~1, data = grizz, weights = newgrizz$weightvec) 

# List of what the function's arguments do:
# ids = formula/dataframe identifying cluster ids from largest level to smallest level (-0) and -1 is a 
# formula for no clusters. 
# data = dataframe
# weights = out weights to be used in the data. Select weights from anesrake object we created 
# earlier and use sigil to select numeric column title "weightvec" from it. 

summary(wt_grizz_df) # Need to use this and other functions from survey package with weighted data. 

# See weights 

sup_unwt <-  wpct(grizz$griz_support) 
sup_wt  <-  wpct(grizz$griz_support, newgrizz$weightvec) # See ** 
sup_tab <- data.frame(sup_unwt, sup_wt)
sup_tab

# ** You can weight your table by providing a vector of weights in wpct function, which I do
# by calling the weightvector from newgrizz. 

table(grizz$griz_support)/length(grizz$X)*100

# Here I produce a table to check that wpct calculated the unweighted proportion correctly. 

grizz$weights <- newgrizz$weightvec

# Here I create a csv file for my weights. 

write.csv(print(newgrizz), "GRIZZ_WT_USE.csv")
```

##Pooled Factor Analysis
The factanal command uses the maximum likelihood to identify the loadings of different components
```{r, echo=FALSE}
grizz_factor<-cbind(grizz[,47:61])
names(grizz_factor)<-c("survival", "forests", "SafetyThreat", "LivelihoodThreat", "Tourism",
                       "Feds", "ThreatPP", "OtherSpecies", "Extinction", "LocalControl", "Ranchers",
                       "OutdoorRec", "Ag", "Urban", "Rural")
grizz_comp<-factanal(grizz_factor, 5, scores="Bartlett")
grizz_comp

#Plotting the relationship of the factors
load<-grizz_comp$loadings[,1:2]
# plot(load, type="n", main="Factors Influencing Support for Grizzly Reintroduction") # set up plot 
# text(load,labels=names(grizz_factor),cex=1.3)
# 
# fa.diagram(grizz_comp$loadings)

#grizz<-cbind(grizz, grizz_comp$scores)
```

##Values
```{r, echo=FALSE}
grizz_values<-read.csv("kunisurvey_coded_march8.csv", header=T)
values<-cbind(grizz_values[,36:44], grizz_values[,46:54])

calc_values<-function(df, v1, v2, v3){
  dat<-rowSums(df[,c(v1, v2, v3)], na.rm=TRUE)
}

altruism<-calc_values(values, v1="male_altruism1", v2="male_altruism2", v3="male_altruism3")
egoism<-calc_values(values, v1="male_ego1", v2="male_ego2", v3="male_ego3")
biospherism<-calc_values(values, v1="male_bio1", v2="male_bio2", v3="male_bio3")
male_sum<-data.frame(altruism, egoism, biospherism)

altruism<-calc_values(values, v1="female_altruism1", v2="female_altruism2", v3="female_altruism3")
egoism<-calc_values(values, v1="female_ego1", v2="female_ego2", v3="female_ego3")
biospherism<-calc_values(values, v1="female_bio1", v2="female_bio2", v3="female_bio3")
female_sum<-data.frame(altruism, egoism, biospherism)

values_tot<-female_sum+male_sum
values_adj<-values_tot/3


write.csv(grizz,file="grizz_calc_values_10.25.csv")
```

Breaking out the biospherism values:
```{r, echo=FALSE}
bio1<-rowSums(values[,c("male_bio1", "female_bio1")], na.rm=TRUE)
bio2<-rowSums(values[,c("male_bio2", "female_bio2")], na.rm=TRUE)
bio3<-rowSums(values[,c("male_bio3", "female_bio3")], na.rm=TRUE)
```

##Regression
```{r, echo=FALSE}
sup_ols_lm_wts <- lm(griz_support ~ 
                       as.factor(college) # TRUE = College degree; FALSE = No college degree.
                     + as.factor(rural) # TRUE = Rural; FALSE = Not Rural. 
                     + as.factor(gender) # 1 = Male; 2 = Female. 
                     + age
                     + hhincome 
                     + as.factor(latino) # 1 = Latino; 2 = Not Latino. 
                     + ideology # Increase corresponds to being more conservative
                     + as.factor(partyid) # 1 = Rep.; 2 = Dem.; 3 = Ind.; 4 = Other. 
                     #+ as.factor(votechoice) # 1 = Trump; 2 = Clinton; 3 = Other; 4 = Didn't Vote. 
                     + as.factor(species_griz) # 1 = Yes; 2 = No; 3 = Don't Know. 
                     + Factor1
                     + Factor2
                     + Factor3
                     + Factor4
                     + Factor5
                     + altruism
                     + biospherism
                     + egoism
                     + Q66 # Unit increase corresponds to increased likelihood to discontinue.
                     + as.factor(natpark_ca) # 1 = Yes; 2 = No; 3 = Don't Know.
                     + as.factor(gtreat1num), # 1 = Treated; 0 = Control.
                     weights = weights,
                     data = grizz)

summary(sup_ols_lm_wts)
#xtable(sup_ols_lm_wts, out="regression.htm")
#write.table(summary(sup_ols_lm_wts), file="wtRegresseion.txt", sep=",", quote=FALSE, row.names=F)
#stargazer(summary(sup_ols_lm_wts), out="regression.htm")
```

With broken out biospherism values:
```{r, echo=FALSE}

sup_ols_lm_wts_bio <- lm(griz_support ~ 
                       as.factor(college) # TRUE = College degree; FALSE = No college degree.
                     + as.factor(rural) # TRUE = Rural; FALSE = Not Rural. 
                     + as.factor(gender) # 1 = Male; 2 = Female. 
                     + age
                     + hhincome 
                     + as.factor(latino) # 1 = Latino; 2 = Not Latino. 
                     + ideology # Increase corresponds to being more conservative
                     + as.factor(partyid) # 1 = Rep.; 2 = Dem.; 3 = Ind.; 4 = Other. 
                     #+ as.factor(votechoice) # 1 = Trump; 2 = Clinton; 3 = Other; 4 = Didn't Vote. 
                     + as.factor(species_griz) # 1 = Yes; 2 = No; 3 = Don't Know. 
                     + Factor1
                     + Factor2
                     + Factor3
                     + Factor4
                     + Factor5
                     + altruism
                     + bio1
                     + bio2
                     + bio3
                     + egoism
                     + Q66 # Unit increase corresponds to increased likelihood to discontinue.
                     + as.factor(natpark_ca) # 1 = Yes; 2 = No; 3 = Don't Know.
                     + as.factor(gtreat1num), # 1 = Treated; 0 = Control.
                     weights = weights,
                     data = grizz)

summary(sup_ols_lm_wts_bio)

```