---
title: "Grizzly Factor Analysis and PCA"
author: "Elizabeth Hiroyasu"
date: "January 25, 2018"
output: pdf_document
---

Factor analysis conducted to assess how people think about grizzly bear reintroduction and what types of variables those attitudes map onto.
```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
library(psych)
library(FactoMineR)
library(nFactors)

grizz <- read.csv("./data/processed/GRIZZLY_DATA_USE.csv", header = T)

grizz_factor<-cbind(grizz[,c(which(colnames(grizz)=="Q72_16"):which(colnames(grizz)=="Q72_32"))])
names(grizz_factor)<-c("survival", "forests", "SafetyThreat", "LivelihoodThreat", "Tourism",
                       "Feds", "ThreatPP", "OtherSpecies", "Extinction", "LocalControl",
                       "Ranchers", "OutdoorRec", "Ag", "Urban", "Rural")
```

We are specifically interested in people's attitudes toward the reintroduction of grizzly bears. Respondents were asked to respond with how much they agreed or disagreed on a 7-point likert scale with the following questions:

1.  People have a responsibility to ensure the survival of grizzly bears.
2.  Grizzly bear reintroduction would help make California forests healthier.
3.  Grizzly bear reintroduction would pose a threat to my safety.*
4.  Grizzly bear reintroduction would pose a threat to my livelihood.*
5.  Grizzly bear reintroduction would benefit the California economy by increasing tourism.
6.  Grizzly bear reintroduction would lead to an increased role for the federal government.
7.  Grizzly bear reintroduction would threaten property rights on private lands.
8.  Grizzly bear reintroduction would benefit other species.
9.  Grizzly bear reintroduction would help prevent their extinction.
10.  Grizzly bear reintroduction would reduce local control over public lands.
11.  Grizzly bear reintroduction would negatively impact ranchers.
12.  Grizzly bear reintroduction would benefit outdoor recreation.
13.  Grizzly bear reintroduction would harm agricultural producers.
14.  Grizzly bear reintroduction would benefit urban residents.
15.  Grizzly bear reintroduction would benefit rural residents.


##Total PCA
This analysis includes personal threat questions
note:prcomp is preferred to princomp

```{r, echo=FALSE}
grizz.pca1<-prcomp(grizz_factor)
summary(grizz.pca1)
plot(grizz.pca1, type="l")
biplot(grizz.pca1)
```

PCA loadings
```{r, echo=FALSE}
loadings<-grizz.pca1$rotation
loadings
```

head of PCA scores
```{r, echo=FALSE}
scores<-grizz.pca1$x
head(scores)
```

##Suppose we want to determine ahead of time the number of factors to extract:
```{r, echo=FALSE}
ev <- eigen(cor(grizz_factor))# get eigenvalues
ap <- parallel(subject=nrow(grizz_factor),var=ncol(grizz_factor), rep=100,cent=.05)
nS <- nScree(x=ev$values, aparallel=ap$eigen$qevpea)
plotnScree(nS)
```


##Pooled PCA:
We can run the principal components analysis dropping the questions that include language of personal threats.
The pooled PCA suggests that the questions collapse onto two main factors
```{r, echo=FALSE}
#dropping out personal threat questions
grizz_factor2<-subset(grizz_factor, select=-c(SafetyThreat, LivelihoodThreat))
```

###Parallel analysis
this tells us that we can preserve two factors
```{r, echo=FALSE}
ev <- eigen(cor(grizz_factor2))# get eigenvalues
ap <- parallel(subject=nrow(grizz_factor2),var=ncol(grizz_factor2), rep=100,cent=.05)
nS <- nScree(x=ev$values, aparallel=ap$eigen$qevpea)
plotnScree(nS)
```


###PCA
using the principal function from the psych package
```{r, echo=FALSE}
grizz.pca2.principal<-principal(grizz_factor2, rotate="varimax", nfactors=2)
plot(grizz.pca2, type="l")
biplot(grizz.pca2)
```

head of PCA scores
```{r, echo=FALSE}
scores<-data.frame(grizz.pca2.principal$scores)
head(scores)
grizz<-cbind(grizz, scores)
#write.csv(grizz, file="./data/processed/grizz_withPCAScores_060718.csv")
```

Looking at Cronbach's alpha using the components from the pooled PCA
```{r}
grizz_component1<-subset(grizz_factor2, select=c("survival", "forests", "Tourism","OtherSpecies","Extinction","OutdoorRec","Urban","Rural"))
psych::alpha(grizz_component1)

grizz_component2<-subset(grizz_factor2, select=c("Feds", "ThreatPP", "Ranchers", "Ag", "LocalControl"))
psych::alpha(grizz_component2)
```

##################################################################################################

##Using the FactoMineR package for more exploratory principal components analysis
```{r}
grizz.pca2.2<-PCA(grizz_factor2)
summary(grizz.pca2.2)
scores<-grizz.pca2.2$ind$coord
```

##Pooled Factor Analysis including personal threat questions
From here forward, we will be dropping the personal threat questions from the analysis.

The factanal command uses the maximum likelihood to identify the loadings of different components. We used an a priori guess of five factors to load on.
```{r, echo=FALSE}
grizz_comp<-factanal(grizz_factor2, 5, scores="Bartlett")
grizz_comp

#Plotting the relationship of the factors
fa.diagram(grizz_comp$loadings)

#grizz<-cbind(grizz, grizz_comp$scores)
```


This suggests that the 15 questions collapse into two factors. We can then rerun the exploratory factor analysis, specifying two factors.
```{r, echo=FALSE}
grizz_comp2<-factanal(grizz_factor2, 2, scores="Bartlett")
grizz_comp2

fa.diagram(grizz_comp2$loadings)
```


##Pooled NMDS
currently no convergence
Probably because there isn't a variable that binds the questions together. We should have N objects measured on p numeric variables. In this case, we just have p numeric variables, without the objects.
```{r, echo=FALSE}
# library(vegan)
# library(MASS)
# grizz.nmds<-metaMDS(grizz_factor, distance="bray", k=3, trymax=20, noshare=0.1)
# grizz.nmds.2<-metaMDS(grizz_factor, previous.best=grizz.nmds)
# plot(grizz.nmds, display=c("sites", "species"), choice=c(1,2), type="t")
```

